Supplementary Material for ``Provisional Title''

### 2021-05-07 draft/draft.tex
started introduction, added some supplementary figures, coalescence.py ready for simulations with simple, complex and intermediate primary resource

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{units} % nice fractions
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=3cm]{geometry} % margins
\usepackage{lineno} % line numbers

@@

@@

This is an example cite \cite{Vetrovsky2013,nloptr}.
This is how you refer to \figref{fig1}.
This is how you refer to \figref[a]{fig1}.
This is how you refer to \tableref{fig1}.
This is how you refer to the section \nameref{methods:sim} of the \nameref{methods}.
This is a cleaner way to refer to the \methodsref{methods:sim}.

@@

% references
\clearpage
\bibliographystyle{mystyle}
%\bibliographystyle{unsrt}
\bibliography{refs}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=10cm,keepaspectratio]{figs/fig1.pdf}
\caption{\textbf{Overview of the experimental protocol.}
\textbf{a.}~Environmental samples collected from eight different locations were used
to inoculate our communities.
\textbf{b.}~Communities were stabilized in serial batch culture bioreactors
\cite{Goldford2018} in minimal synthetic media with glutamine or citrate as the
only supplied carbon source.
\textbf{c.}~Communities were plated in minimal media agar plates and the most abundant
species (the ``dominants") from each community were isolated. We refer to the set of
sub-dominant species as the ``cohorts".
\textbf{d.}~Rank-frequency distributions of all eight communities stabilized in either
glutamine (red) or citrate (blue), sequenced at a depth of $10^{-4}$ reads.
Three biological replicates per community are shown.
Community compositions are skewed and long-tailed.
\textbf{e.}~Our hypothesis is that ecological co-selection can take place from the top-down,
i.e. the dominant co-selecting the cohort, or from the bottom-up, i.e. the cohort co-selecting
the dominant. Both forms of co-selection can be positive (recruitment) or negative
(antagonism).
\textbf{f.}~Illustration of the protocol of our coalescence experiments. All pairs of
communities were inoculated into fresh minimal media supplemented with the same carbon
source where communities had been previously stabilized. The coalesced (C) and original
resident (R) and invasive (I) communities were then serially diluted and allowed to grow
for seven additional transfers.}
\label{fig1}
\end{figure}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=15cm,keepaspectratio]{figs/fig2.pdf}
\caption{\textbf{Co-selection in microbial community coalescence.}
\textbf{a.}~Coalescence outcomes are quantified by the relative Bray-Curtis similarity
($Q$) between the coalesced and invasive communities (\methodsref{methods:metrics}).
These outcomes are predicted by the pairwise competition between the invasive and
resident dominant
species ($R^2=0.15$ for glutamine and $R^2=0.57$ for citrate). This is consistent with a scenario of
top-down positive co-selection where dominants recruit their cohorts for the final
coalesced community.
\textbf{b.}~We represent the frequency reached by
the invasive dominant species when they invade the resident communities on their own versus
when they are in the company of their cohort.
Three scenarios are possible: green and red
shaded areas represent limit cases of positive (recruitment) or negative (antagonism)
bottom-up co-selection, gray area corresponds to situations where invasive dominant
species can invade with equal success regardless of the presence of their cohorts.
Data shows that positive co-selection is common, whereas antagonistic
co-selection is rare in our experiments.
\textbf{c-d.} Simulations of community coalescence with a consumer-resource model
are able to capture these trends.}
\label{fig2}
\end{figure}

@@

>>>>

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{units} % nice fractions
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\zxrsetup{toltxlabel}
\zexternaldocument*{supp}
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=3cm]{geometry} % margins
\usepackage{lineno} % line numbers
\usepackage{textcomp} % special characters (some references break if this isn't loaded?)

@@

\section*{Test section}

This is an example cite \cite{Luo2020,Vass2021}.
This is how you refer to \figref{fig1}.
This is how you refer to \figref[a]{fig1}.
This is how you refer to \tableref{fig1}.
This is how you refer to \figref{figS1}.
This is how you refer to the section \nameref{methods:sim} of the \nameref{methods}.
This is a cleaner way to refer to the \methodsref{methods:sim}.

@@

Microbial communities often invade one another.
This has been observed, for instance, 
in river courses where terrestrial microbial communities mix with aquatic microorganisms
\cite{Luo2020,Vass2021}
or in soil communities being invaded by aerially dispersed bacteria and funghi \cite{Evans2019}.
The human digestive system can even get invaded several times a day by the microbial consortia
that reside on the ingested food.
Ecologists have long contemplated the idea that interactions between multiple co-invading species
can produce correlated invasional outcomes
\cite{Gilpin1994,Grosholz2005,Green2011,Livingston2013,Rillig2015,OLoughlin2017}.
One example is the hypothesis known as \textit{invasional meltdown}, which proposes that
positive interactions between co-invading species can enhance their invasive success and
facilitate future invasions \cite{Simberloff1999,Simberloff2006,Gurevitch2006}.

The phenomenon by which entire microbiomes invade one another has been termed
\textit{community coalescence} \cite{Rillig2015}.
In spite of its clear potential importance, the role of coalescence in microbiome assembly is
only beginning to be addressed and little is known about the mechanisms that govern it and its
potential implications.
Early mathematical models of community-community invasions \cite{Gilpin1994,Toquenaga1997},
as well as more recent work including environmental construction (e.g., consumer-resource models)
\cite{Tikhonov2016,Tikhonov2017,Lechon2021}, suggest that high-order invasion effects are common
during community coalescence. Communities that have a previous history of coexistence may exhibit an
emergent ``cohesiveness'' which produces correlated invasional outcomes among species from the
same community \cite{Livingston2013,Sierocinski2017}.
The situation where ecological partners in the invading community recruit one another into the final
coalesced community has been called \textit{ecological co-selection} \cite{Rillig2017,Sierocinski2017}.

@@

% references
\clearpage
\bibliographystyle{mystyle}
\bibliography{refs}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=10cm,keepaspectratio]{figs/fig1.pdf}
\caption{\textbf{Overview of the experimental protocol.}
\textbf{a.}~Environmental samples collected from eight different locations were used
to inoculate our communities.
\textbf{b.}~Communities were stabilized in serial batch culture bioreactors
\cite{Goldford2018} in minimal synthetic media with glutamine or citrate as the
only supplied carbon source.
\textbf{c.}~Communities were plated in minimal media agar plates and the most abundant
species (the ``dominants") from each community were isolated. We refer to the set of
sub-dominant species as the ``cohorts".
\textbf{d.}~Rank-frequency distributions of all eight communities stabilized in either
glutamine (red) or citrate (blue), sequenced at a depth of $10^{-4}$ reads.
Three biological replicates per community are shown.
Community compositions are skewed and long-tailed.
\textbf{e.}~Our hypothesis is that ecological co-selection can take place from the top-down,
i.e. the dominant co-selecting the cohort, or from the bottom-up, i.e. the cohort co-selecting
the dominant. Both forms of co-selection can be positive (recruitment) or negative
(antagonism).
\textbf{f.}~Illustration of the protocol of our coalescence experiments. All pairs of
communities were inoculated into fresh minimal media supplemented with the same carbon
source where communities had been previously stabilized. The coalesced (C) and original
resident (R) and invasive (I) communities were then serially diluted and allowed to grow
for seven additional transfers.}
\label{fig1}
\end{figure}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=15cm,keepaspectratio]{figs/fig2.pdf}
\caption{\textbf{Co-selection in microbial community coalescence.}
\textbf{a.}~Coalescence outcomes are quantified by the relative Bray-Curtis similarity
($Q$) between the coalesced and invasive communities (\methodsref{methods:metrics}).
These outcomes are predicted by the pairwise competition between the invasive and
resident dominant
species ($R^2=0.15$ for glutamine and $R^2=0.57$ for citrate). This is consistent with a scenario of
top-down positive co-selection where dominants recruit their cohorts for the final
coalesced community. Two biological replicates per experiment are plotted individually.
\textbf{b.}~We represent the frequency reached by
the invasive dominant species when they invade the resident communities on their own versus
when they are in the company of their cohort.
Three scenarios are possible: green and red
shaded areas represent limit cases of positive (recruitment) or negative (antagonism)
bottom-up co-selection, gray area corresponds to situations where invasive dominant
species can invade with equal success regardless of the presence of their cohorts.
Data shows that positive co-selection is common, whereas antagonistic
co-selection is rare in our experiments.
Two biological replicates per experiment are plotted individually.
\textbf{c-d.} Simulations of community coalescence with a consumer-resource model
are able to capture these trends.}
\label{fig2}
\end{figure}

@@

\section*{Supplementary Figures}

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=15cm,keepaspectratio]{figs/figS1.pdf}
\caption{\textbf{Community compositions after seven additional transfers without coalescence.}
Each color of the pie plots corresponds to a different exact sequence variant
(\methodsref{methods:sequencing}).
Replicate 2 of community 1 from glutamine,
as well as replicate 2 of community 8 from citrate (highlighted) were removed based on their
dissimilarity to the other two replicates (details in code for data analysis, see
\nameref{datacode}).
Communities clustered in dashed boxes shared the same dominant species as revealed by
sequencing data.
For communities enclosed in red boxes, sequencing data showed that the species isolated by plating
was not detectable in the community after seven additional transfers (i.e. the dominant was
incorrectly identified) and were therefore excluded from downstream analyses.}
\label{figS1}
\end{figure}

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=12.5cm,keepaspectratio]{figs/figS2.pdf}
\caption{\textbf{Alternative metrics of community distance.}
Quantifying coalescence outcomes using different metrics of community similarity
(\methodsref{methods:metrics}) gives similar results to those shown in \figref[a]{fig2}.
Metrics that account for the relative species abundances (\textbf{a.} Jensen-Shannon similarity or
\figref[a]{fig2} Bray-Curtis similarity) yield higher correlations than less quantitative metrics that only
account for species presence/absence (\textbf{b.} Jaccard similarity or \textbf{c.} Fraction of endemic
invasive species persisting in the coalesced community).}
\label{figS2}
\end{figure}

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=10cm,keepaspectratio]{figs/figS3.pdf}
\caption{\textbf{Dominant species have limited effects on coalescence outcomes.}
We repeated the analyses shown in \figref[a]{fig2} and \figref{figS2}, but this time we removed the
dominants from the compositional data prior to quantifying community distances.
The trends observed before are maintained.}
\label{figS3}
\end{figure}

### 2021-05-07 draft/supp.tex
started introduction, added some supplementary figures, coalescence.py ready for simulations with simple, complex and intermediate primary resource

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{mathtools} % nice matrices
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage{xcolor} % text color
\usepackage[margin=3cm]{geometry} % margins
\usepackage[nottoc,numbib]{tocbibind} % add references to table of contents

@@

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\addto{\captionsenglish}{\renewcommand{\bibname}{Supplementary References}}
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\renewcommand{\thefigure}{\textbf{S\arabic{figure}}} % preceed the Suppl Mat figures with S
\renewcommand{\thetable}{\textbf{S\arabic{table}}} % preceed the Suppl Mat tables with S
\renewcommand{\theequation}{S\arabic{equation}} % preceed the Suppl Mat equations with S
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Fig.}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions
\hypersetup{citecolor={blue}}

@@

\maketitle
\tableofcontents
\newpage

\section{Supplementary Methods}\label{supp-methods}

\subsection{Data processing}\label{supp-methods:data-proc}

\subsubsection{Examining community composition}
\label{supp-methods:data-proc:community-composition}

We first analyze the composition of the communities assembled in each carbon
source by examining the relative abundances of the exact sequence variants
(ESVs) obtained from 16S rRNA gene sequencing. We define a vector $\mathbf{s}
= \left( s_1, s_2, \cdots, s_M \right)$ such that $s_i$ represents the relative
abundance
of the $i$-th ESV, being $M$ the total number of unique ESVs. We assembled
a total of 16 communities, 8 in glutamine
and 8 in citrate (see Methods \mr), with 3 biological replicates of each. We
start
by comparing the composition of replicate communities in terms of their ESV
relative abundances. To do this, we measure the distance $d$ between two
communities (with ESV abundances $\mathbf{s}$ and $\mathbf{s}'$ respectively) as

\begin{equation}
d \left( \mathbf{s}, \mathbf{s}' \right) =
\sqrt{\sum_i{\left(s_i - s'_i\right)^2}}
\label{supp-eqn:esvdist}
\end{equation}
%
We compute all the distances across replicates of the same community, obtaining
a set of 48 distances (16 communities $\times$ 3 pairwise distances between the
three replicates) that we will denote as $\left\{d\right\}$

\begin{equation} 
\left\{d\right\} = 
\left\{ d\left(\mathrm{community} \; i \; \mathrm{replicate} \; j,
\mathrm{community} \; i \; \mathrm{replicate} \; j' \right) \;
\text{for all $i$, $j$, $j' \neq j$} \right\}
\end{equation}
%
We then define a threshold $d_T$ as

\begin{equation}
d_T = \mathrm{Q}4 \left( \left\{d\right\} \right) + 
\mathrm{IQR} \left( \left\{d\right\} \right)
\label{supp-eqn:threshold}
\end{equation}
%
Where Q4 and IQR represent the fourth quantile and the interquartile range
respectively. Note that this threshold is analogous to the one used in
standard boxplots to identify outliers.

Finally, we discard the samples that are at a distance larger than $d_T$ from
all other replicates of the same community. These are replicate 2 of community
1 in glutamine and replicate 2 of community 8 in citrate (see Figure S1 \mr).
The remaining replicates were used in further analyses. We took the average
ESV abundance across replicates when indicated.

\subsubsection{Special cases}\label{supp-methods:data-proc:special}

Blablabla.

\subsection{Inferring species abundance from sequencing data}
\label{supp-methods:species-from-seq}

Consider a community of $N$ species with relative abundances represented
by the components of a vector $\mathbf{x} = \left( x_1, x_2, \dotsc,
x_N\right)$ such that $x_i$ is the relative abundance of the $i$-th species.
Naturally, the conditions
%
\begin{equation}
\sum_i x_i = 1
\label{supp-eqn:normx_1}
\end{equation}
%
and
%
\begin{equation}
0 \leq x_i \leq 1 \; \textrm{for all} \; i
\label{supp-eqn:normx_2}
\end{equation}
%
are satisfied.

Performing 16S rRNA gene sequencing on such a community yields a list of
sequences (exact sequence variants or ESVs) and their respective abundance.
Normalizing by the total number of sequences, we can obtain a vector $\mathbf{s}
= \left( s_1, s_2, \dotsc \right)$ that also satisfies the normalization
conditions of equations \ref{supp-eqn:normx_1} and \ref{supp-eqn:normx_2}.
Only if there are as many ESVs as species does $\mathbf{s}$ have length $N$
(the same as $\mathbf{x}$), but in the most general case it is possible that
\textit{a)} multiple species share a same ESV and/or \textit{b)} species carrying
multiple copies
of the 16S rRNA gene have different sequences for each copy \cite{Vetrovsky2013}. We denote
the length of $\mathbf{s}$ as $M$, which need not be equal to $N$.

By sequencing each species individually, we can build a $M \times N$ 
matrix $\mathbf{Q}$ such
that the element in the $(i,j)$ position, $q_{ij}$, represents the frequency of
the $i$-th ESV when the $j$-th species is sequenced. $\mathbf{Q}$ can be used
to determine the fraction of sequences of a given ESV that are obtained from
sequencing a community with any arbitrary composition. In other words,
$\mathbf{Q}$ maps $\mathbf{x}$ to $\mathbf{s}$:

\begin{equation}
\mathbf{Q} \cdot \mathbf{x} = \mathbf{s}
\label{supp-eqn:Qxs}
\end{equation}

Note that equation \ref{supp-eqn:Qxs} is only true if all species have the same
number of copies of the 16S rRNA gene. Otherwise, species with high copy numbers
will yield more sequences, thus leading to an over-representation and vice-versa.
Information on the 16S copy number for every species in the communities would be
required to overcome this limitation, but this is not feasible when dealing with
highly diverse communities. For our purpose in this
work, we will need to assume that the copy numbers of the 16S are relatively
conserved across species and equation \ref{supp-eqn:Qxs} holds. Additionally,
we have sequencing data from a subset of the species in our
communities, but not from all of them. This means that we cannot build a complete
$\mathbf{Q}$ matrix. To address this, we first identify every ESV obtained
from community sequencing that we cannot map back to (at least) one of the
species for which we have single-species sequencing information. We then
assume that each
of those ESVs maps uniquely to a single species. This gives us a $\mathbf{Q}$
matrix of the form:

\begin{equation}
\mathbf{Q} = \; \; \; \; \; \begin{matrix}
 &

\begin{matrix}
\; \; \;
\rotatebox{90}{\scriptsize species 1} \; \; \; & 
\rotatebox{90}{\scriptsize species 2} \; \; \; &
\rotatebox{90}{\scriptsize species 3} \; \; &
\cdots \; \; &
\rotatebox{90}{\scriptsize species $N-2$} \! \! & 
\rotatebox{90}{\scriptsize species $N-1$} \! \! &
\rotatebox{90}{\scriptsize species $N$}
\end{matrix}

\\

\begin{matrix*}[r]
\rotatebox{0}{\scriptsize sequence 1} \\
\rotatebox{0}{\scriptsize sequence 2} \\
\rotatebox{0}{\scriptsize sequence 3} \\
\vdots \; \; \; \\
\rotatebox{0}{\scriptsize sequence $M-2$} \\ 
\rotatebox{0}{\scriptsize sequence $M-1$} \\
\rotatebox{0}{\scriptsize sequence $M$}
\end{matrix*}

&

\begin{pmatrix}
q_{11} & q_{12} & q_{13} &  & 0 & 0 & 0 \\
q_{21} & q_{22} & q_{23} & \cdots & 0 & 0 & 0 \\
q_{31} & q_{32} & q_{33} &  & 0 & 0 & 0 \\
 & \vdots &  & \ddots  &  &  &  \\
0 & 0 & 0 &  & 1 & 0 & 0 \\ 
0 & 0 & 0 &  & 0 & 1 & 0 \\ 
0 & 0 & 0 &  & 0 & 0 & 1 
\end{pmatrix}

\end{matrix}
\label{supp-eqn:Q_modified}
\end{equation}
%
While a one-to-one mapping between species and ESVs may not be true in all cases,
in practice the ESVs for which we make this assumption have low relative
abundances in the communities, minimizing the impact of any potential artifact.

Having built our $\mathbf{Q}$ matrix according to equation
\ref{supp-eqn:Q_modified}, obtaining species abundances from sequencing data
is relatively straigthforward: we need to find the $\mathbf{x}$ that satisfies
equation \ref{supp-eqn:Qxs} for a given $\mathbf{s}$. However, in some cases
sequencing error can introduce deviations in $\mathbf{s}$ that make it so the
vector $\mathbf{x}$ that solves equation \ref{supp-eqn:Qxs} does not meet the
boundary conditions in equation \ref{supp-eqn:normx_2}, i.e., some $x_i$ may be
negative or greater than 1. These cases are obviously problematic, but we can
avoid them by accounting for potential deviations in $\mathbf{s}$. We do this
by solving the following nonlinear optimization problem: we want to obtain an
estimate of the true community composition
(we will denote the estimate as $\mathbf{\hat{x}}$) so that the product
$\mathbf{Q} \cdot \mathbf{\hat{x}}$ is as close as possible to the $\mathbf{s}$
obtained from sequencing while
having $\mathbf{\hat{x}}$ satisfy the normalization condition in equation
\ref{supp-eqn:normx_1} and the boundary conditions in equation
\ref{supp-eqn:normx_2}. For every possible $\mathbf{\hat{x}}$, we can define a vector
$\boldsymbol{\upvarepsilon} = \boldsymbol{\upvarepsilon} \left( \mathbf{\hat{x}} \right) = 
\left( \varepsilon_1, \varepsilon_2, \cdots, \varepsilon_M \right)$ as

\begin{equation}
\boldsymbol{\upvarepsilon} \left( \mathbf{\hat{x}} \right) =
\mathbf{Q} \cdot \mathbf{\hat{x}} - \mathbf{s}
\end{equation}
%
and a function $f \left( \mathbf{\hat{x}} \right)$ as

\begin{equation}
f \left( \mathbf{\hat{x}} \right) =
\left| \boldsymbol{\upvarepsilon} \left( \mathbf{\hat{x}} \right) \right| ^2 =
\sum_i \varepsilon_i^2 \left( \mathbf{\hat{x}} \right)
\end{equation}
%
We can also define a function $h \left( \mathbf{\hat{x}} \right)$ as

\begin{equation}
h \left( \mathbf{\hat{x}} \right) = 1 - \sum_i\hat{x}_i
\end{equation}
%
The best estimation for the species composition of a community will be the
$\mathbf{\hat{x}}$ that minimizes $f \left( \mathbf{\hat{x}} \right)$ while
satisfying the normalization condition (eq. \ref{supp-eqn:normx_1}) and the
boundary conditions (eq. \ref{supp-eqn:normx_2}). We solve
this problem using the \texttt{nloptr} package for R \cite{nloptr}, passing it
$f \left( \mathbf{\hat{x}} \right)$ as the function to minimize,
$h \left( \mathbf{\hat{x}} \right) = 0$ as an equality constraint,
and 0 and 1
as the lower and upper bounds for the entries of $\mathbf{\hat{x}}$, and using
the augmented Lagrangian algorithm \cite{auglag1,auglag2}. Further
details can be found in the code for the analysis (see section 
\nameref{supp-datacode} of this Supplementary Material).

\section{Data \& code availability}
\label{supp-datacode}

Data and code for the analyses in this article, as well as code for the
\texttt{community-simulator} package updated with the new functionalities,
is available at \url{https://github.com/jdiazc9/coalescence}.

\section{Supplementary Results}\label{supp-results}

Blabla

\subsection{Example subsection}
\label{supp-results:example}

Blabla

\section{Supplementay Figures}\label{supp-figs}

Blablabla.

>>>>

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\zxrsetup{toltxlabel}
\zexternaldocument*{draft}
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=3cm]{geometry} % margins
\usepackage[nottoc,numbib]{tocbibind} % add references to table of contents

@@

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\addto{\captionsenglish}{\renewcommand{\bibname}{Supplementary References}}
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\renewcommand{\thefigure}{\textbf{S\arabic{figure}}} % preceed the Suppl Mat figures with S
\renewcommand{\thetable}{\textbf{S\arabic{table}}} % preceed the Suppl Mat tables with S
\renewcommand{\theequation}{S\arabic{equation}} % preceed the Suppl Mat equations with S
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Fig.}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions
\hypersetup{citecolor={blue}}
\definecolor{lightgray}{gray}{0.9}
\newcommand{\figref}[2][]{%
  \hyperref[{#2}]{%
    Figure~\ref*{#2}%
    \ifx\\#1\\%
    \else
      #1%
    \fi
  }%
}
\newcommand{\tableref}[1]{%
  \hyperref[{#1}]{%
   Table~\ref*{#1}%
  }%
}
\newcommand{\methodsref}[1]{%
  \hyperref[{#1}]{%
   Methods:~\nameref*{#1}%
  }%
}

@@

\maketitle
\newpage

\section*{Supplementay Figures}\label{supp-figs}

This is how you reference something from the draft...

\begin{figure}[!h]
\centering
\includegraphics[width=15cm,keepaspectratio]{figs/figS1.pdf}
\caption{\textbf{Community compositions after seven additional transfers without coalescence.}
Each color of the pie plots corresponds to a different exact sequence variant
(\zref{fig1}).}
\label{figS1}
\end{figure}
